Draft

### 2017-09-08 Thesis/Draft.Rmd
Add descriptions to Appendix

The thesis at hand joins the ranks of a large literature concerned with the general problem of sequential sampling. One is familiar with the idea of "batch sampling", where the entire sample is gathered first and analyzed only after the last sample has become available. However, as noted in Chernoff (1958), many situations arise in which one would like to gather feedback from samples as they are drawn so as to influence the choice of the next sample. The two fundamental characteristics of the latter sampling process are that it is sequential and adaptive. Since observations are drawn in sequence, one has a chance to observe the observations drawn so far in order to adapt the decision of what sample one would like to observe next.

This makes sequential sampling especially relevant in situations where there is a inherent cost to drawing an observation. The classical example that of clinical trials in which a researcher tests two drugs on patients to observe which one performs better. Not only is it difficult to gather patients, but also one would like to adapt the sampling such that as soon as possible more patients receive the better drug so that more patients are cured. As noted in Cappé et al. (2013), this example dates back to Thompson (1933, 1935). Such problems are known as multi-armed bandit problems in which the sampling is adapted so as to minimize the cumulative regret: In the case of the drug tests, one would like to sample the two drugs in a way to find the best drug as soon as possible and then to test this drug more and more often and more often than the worse drug. This would lead to cumulatively more healthy patients then when both drugs are tested uniformly and thus minimize the regret from treating a patient with an ineffective drug.

Instead of drugs, can consider two or more unknown distributions (which we call arms in the multi-armed bandit problem). These distributions have different means, and one can draw samples sequentially from the distributions in order to estimate their means. One would like to play the distribution with the highest mean to minimize the cumulative regret measure. How can we organize the sampling in order to achieve this goal?

While Gittins (1979) presents an Bayesian-optimal solution, the Upper-Confidence Bound algorithms pioneered by Agrawal (1995) and Auer, Cesa-Bianchi and Fischer (2002) lead to a blooming literature on multi-armed bandit problems that has been picked up to solve modern problems in website and advertising optimization, where advertisers try to allocate their advertising optimally so as to reduce regret from showing poorly converting advertisements. Several algorithms as the KL-UCB algorithm based on Kullback-Leibler divergence upper confidence intervals (Cappé et al., 2013) or the Bayes-UCB algorithm (Kaufmann et al., 2012)  have followed and been shown to be optimal as they reach the lower bound introduced in Lai and Robbins (1985) and Burnetas and Katehakis (1996).

These algorithms have in common that they compute after any sample an index for each of the unknown distributions. This index is based solely on the past (observed) data, and then used to pick the distribution from which the next observation should be drawn. Using these indexes, the algorithms can sequentially sample and adapt the future strategies to observations as they are gathered.

Other algorithms based on indices can be used on related problems. Instead of minimizing cumulative regret, it might be more interesting to find the best arm, that is, the distribution with the largest mean, with as few samples as possible and with confidence as large as possible. This is particularly useful when there is not necessarily a cost from initially picking a lower mean, but when drawing any sample is costly, and one needs to explore the distributions as efficiently as possible. The need for strategies other than uniform sampling arises: Do not continue to sample distributions that we are confident about. As such, the best arm identification problem is very similar to the problem faced in Bayesian Optimization: Training large neural networks is costly in terms of time and computation, and so the parameter space needs to be explored with as few samples as possible. (Zhang et al., 2017; Brochu et al., 2010).

The problem of best arm identification is very related to statistical hypothesis testing as noted by Chernoff (1958) since in the end, one has to make a decision about how the means of distributions relate to each other. The main difference of course being that the sampling procedure becomes integrated into the evaluation of the distributions. Given the focus on exploration, the problem is usually described with a loss function different from cumulative regret. Audibert et al. (2010) introduce a soft measure called *simple regret* based on which they describe UCB policies. Other authors investigate hard, binary measures that are equal to 1 when the best arm has not been identified (Russo, 2016; Garivier and Kaufmann, 2016). These binary measures correspond to the probability of the identified arm being the best, and thus inherently related to hypothesis testing. The problem of best arm identification can readily be extended to the so called Top-$m$ problem of identifying the best $m$ of $K$ arms (Kaufmann and Kalyanakrishnan, 2013; Kaufmann et al. 2016).

In this thesis, we consider the thresholding bandit problem. Instead of identifying the best $m$ out of $K$ arms, one aims to find all out of $K$ distributions that have a mean larger than a threshold fixed upfront. In the form described in Locatelli et al. (2016), algorithms in the thresholding bandit problem are evaluated against a binary loss function penalizing whenever an arm has not been classified correctly as being above the threshold and thus focus on exploration similar to algorithms employed for best arm identification. Indeed, Locatelli et al. (2016) show how their thresholding algorithm can be used for best arm identification when the mean of the best arm is known upfront.

In the following, we first present the thresholding bandit problem in detail. To then extend upon the algorithm proposed by Locatelli et al. (2016), we first explore how the Kullback-Leibler divergence has been employed in the best arm identification literature by Garivier and Kaufmann (2016) and Kaufmann et al. (2016). We shortly describe two variance based algorithms for the thresholding bandit problem introduced in Mukherjee et al. (2017) and Zhang et al. (2017), before introducing the new Simple Likelihood Ratio algorithm. We end with experiments to examine the performance of different algorithms.

@@

To introduce the thresholding bandit problem, consider a game in which the learner is tasked to find those slot machines within a row of $K$ slot machines in a casino that give a mean payoff larger than a fixed threshold $\tau$, for example $\tau = 0$. The learner cannot pull all slot machines at the same time. Instead, he acquires samples sequentially from one slot machine at a time to learn about their payoff structure. The learner has a budget of $T$ slot machine pulls. After $t-1$ samples, the learner has observed $T_i(t-1)$ samples from slot machine $i$. The payoffs $X_{i,1}, ..., X_{i,T_i(t-1)}$ are drawn from the payoff distribution $\nu_i$ underlying the slot machine $i$. At time $t$, the learner considers the payoffs he observed so far in order to choose the next slot machine he would like to pull. His decision is lead by the goal to classify the slot machines with distribution $\nu_i$ and mean payoff $\mu_i = \mathbb{E}_{X \sim \nu_i}[X]$ after $T$ samples as being above or below the threshold with error probability as small as possible.

The thresholding bandit problem can thus be framed in the wide literature on sequential sampling strategies. While its name is reminiscient of the multi-armed bandit setting, the thresholding bandit problem is more akin to best arm identification and Top-$m$ identification problems due to its focus on exploration: While the classical multi-armed bandit problems are concerned with cumulative regret which increases *during* sampling, best arm identification and thresholding bandits collect loss only once *after* the sampling procedure (compare Garivier and Kaufmann, 2016; Kaufmann et al., 2016; Kaufmann and Kalyanakrishnan, 2013). Furthermore, as there arises an implicit threshold between the means of the $m$-th best arm and the $m+1$-th best arm (Kaufmann et al., 2016; Kaufmann and Kalyanakrishnan, 2013), algorithms for the Top-$m$ identification problem classify arms at least implicitly as being above or below a threshold. The two settings are thus very related.

@@

Consider a setting with a set of $\mathbb{A}=\{1, ..., K\}$ unknown probability distributions $\nu_i, i \in \mathbb{A}$. To stick with the idea of multi-armed bandits, we call those distributions arms, and observations from those arms rewards (Bubeck and Cesa-Bianchi, 2012). The learner is interested in learning about the structure of the arms; in particular, he would like to estimate the means $\mu_1, ..., \mu_K$ by sequentially drawing observations $X_{i,1}, X_{i,2}, ...$ from each arm. However, at time $t$ the learner cannot observe a reward from every distribution but must select a single arm $I_t$ from which he would like to observe the reward $X_{I_t,t}$. An adaptive sampling strategy makes the decision of which arm to pull usually by computing an index $B_i(t)$ for each arm. If $T_i(t)$ is the number of samples observed until time $t$ for arm $i$, then $B_i(t)$ is a function of the past rewards $X_{i,1}, ..., X_{i,T_i(t)}$ observed for arm $i$, and w.l.o.g. $I_t = \arg \min_{i = 1,...K} B_i(t)$. The set of indices can be called strategy (Kaufmann et al., 2012) which makes it clear that the strategy is adapted as more observations are collected. Strategies of this kind are also called index policies.

What index proves to be useful is determined by the loss function with which the decisions of the algorithm are evaluated. While many problems are concerned with finding the arm with the largest mean $\mu_i$, in which algorithms collect loss either cumulatively (Auer and Cesa-Bianchi, 2012) or once at the end of the sampling (Kaufmann et al., 2016), the thresholding bandit problem considers a slightly different loss function. Consider a fixed threshold value $\tau \in \mathbb{R}$ and an epsilon interval $\tau \pm \epsilon$ around the threshold for $\epsilon \geq 0$. As mentioned in the introduction, the learner is tasked with classifying arms as having a mean $\mu_i$ larger or smaller than $\tau$. To describe which arms are above the threshold, we introduce the following set notation (compare Locatelli et al., 2016): Let $\mathcal{S}_{\tau}$ be the set of distributions in $mathbb{A}$ with mean larger than the threshold: $\mathcal{S}_{\tau} := \{i \in \mathbb{A}, \mu_i \geq \tau\}$. Consequently, the complement is given by the arms with mean below the threshold, $\mathcal{S}^C_{\tau} := \{i \in \mathbb{A}, \mu_i < \tau\}$.

The thresholding problem for the learner is now to use a budget of overall $T$ samples (or rounds) to explore the different arms. After this exploration phase, the learner has to return a set of arms that he believes to have a mean above the threshold. The problem can thus be described by the threshold $\tau$ and a precision $\epsilon$. When evaluating the classification of the learner, arms with mean inside of the epsilon interval will not be considered. The learner only has to distinguish the means from the threshold correctly up to the precision $\epsilon$. Thus, the algorithm is not evaluated against $\mathcal{S}_{\tau}$ and $\mathcal{S}^C_{\tau}$, but needs to distinguish $\mathcal{S}_{\tau+\epsilon}$ from $\mathcal{S}^C_{\tau-\epsilon}$. As in Locatelli et al. (2016), whenever we say that "an arm is over the threshold", we mean that the mean of the arm is over the threshold.

@@

$$
\mathcal{L}(T) = \mathbbm{1}{(\hat{\mathcal{S}}_{\tau} \cap \mathcal{S}_{\tau-\epsilon}^C \neq \emptyset) \lor (\hat{\mathcal{S}}_{\tau}^C \cap \mathcal{S}_{\tau+\epsilon} \neq \emptyset)}
$$

In order to perform well, a learner has to identify all arms that are over the threshold as such, and to not claim arms as being over the threshold if they actually are below the threshold. Due to the precision $\epsilon$, the loss is not influenced by arms which are too close to the threshold $\tau$, that is, arms inside of the $\tau \pm \epsilon$ interval. If either $\hat{\mathcal{S}}_{\tau}$ contains an arm below $(\tau - \epsilon)$, or $\hat{\mathcal{S}}^C_{\tau}$ contains an arm above $(\tau + \epsilon)$, a loss of $1$ is accrued. Else, all arms outside the epsilon interval are classified correctly and the loss is equal to $0$. In the case of $\epsilon = 0$, all arms need to be classified correctly to not endure any loss.

@@

which corresponds to the probability of making a wrong classification. Both the probability of not identifying an arm that is actually over the threshold, as well as the probability of accepting an arm as over the threshold when it is not are evaluated. The oracle strategy would evaluate all arms correctly with an error probability of $0$ (Locatelli et al., 2016).

At this point, we have introduced the general thresholding bandit problem setting which we will use throughout the thesis. The setting might be extended by different assumptions to make certain algorithms or bounds feasible. For that reason, Locatelli et al. (2016) assume that the distributions $\nu_i, i \in \mathbb{A}$ are $R$-sub-Gaussian. Under this assumption, the authors are able to prove a lower bound on the largest probability of error of the best algorithm in the setting we have just introduced, as well as an upper bound on the performance of their APT algorithm we will introduce in the following chapter. As in their paper, we define $R$-sub-Gaussian and assume it in the description of their algorithm and bounds:

@@

As noted by Locatelli et al. (2016), this ensures that the mean $\mu_i$ of the distributions $\nu_i$ considered in our problem are finite. $R$-sub-Gaussian distributions furthermore offer concentration inequalities for the concentration of the empirical mean to the true mean, which makes it possible to deduce the bounds presented by the authors and repeated below. The family of $R$-sub-Gaussian distributions encompasses Gaussian distributions of variance $R^2$ for $R \in \mathbb{R}$, as well as all bounded distributions, with the important case of Bernoulli distributions.

@@

Just like the general index policies mentioned in \autoref{sec:Setup}, we see that the APT algorithm adapts its strategy using an index $B_i(t+1)$ for each arm that is based entirely on the observations observed until time $t$. Given the arm with the smallest index is pulled at round $t+1$, we observe that arms that are estimated to be further away from the threshold $\tau$ as measured by the absolute distance to their empirical mean $\hat{\mu}_i$, $\hat{\Delta}_i(t)$, are pulled less often. Intuitively, it is easier to classify arms as below or above $\tau$ when their mean is far away from the threshold. In contrast to the uniform sampling strategy, APT will gather more evidence for arms which appear to be inherently more difficult to classify. Likewise, APT tends to pull arms less often which it already pulled considerably often as measured by $T_i(t)$. If arm $i$ has been pulled more often than arm $j$, then we should in general be more confident in its estimated mean and the corresponding distance to the threshold. If we are confident in $i$, then it is better to explore other arms by drawing an observation from arm $j$. By adapting its sampling strategy in this way during the sequential sampling process, the APT algorithm minimizes the probability of error $\mathbb{E}[\mathcal{L}(T)]$ after $T$ samples have been observed.

The previous paragraph describes how arms close to the threshold are intuitively more difficult to classify. In the index employed by the APT algorithm, the distance between an arm and the threshold is measured by the absolute distance $\hat{\Delta}_i(t)$ between the (estimated) mean of the arm and the threshold $\tau$ (and the epsilon interval). Given that one compares $B_i^{APT}$ with $B_j^{APT}$, one could just as well measure the distance by the squared distance, $\hat{\Delta}_i(t)^2$. When the APT algorithm realizes that an arm has a mean close to the threshold, then this arm requires more attention than an arm far away from the threshold; the confidence in the empirical mean has to be larger. Or the empirical mean has to be better concentrated around the true mean. Otherwise it is easier to make a mistake on this arm. To gain this confidence, the arm requires more samples, and thus a larger share of the overall $T$ samples available. With this intuition, it makes sense to measure the complexity of a thresholding problem: If we knew upfront how complex a given set of arms is, we knew whether we can expect a small or large probability of error given our budget $T$. Also, for example in the best arm identification problem, there exist algorithms which take the complexity of a problem as input. Based on the complexity they then adapt their sampling strategy. See for example the UCB-E algorithm in Audibert et al. (2010) for the best arm identification problem, or the KL-LUCB-E algorithm in Kaufmann and Kalyanakrishnan (2013) as extension to the Top-$m$ problem.

While the APT algorithm does not require the complexity to be known upfront (which would be problematic), it is necessary to define the complexity to deduce the upper bound on its probability of error. Similar to the distance measure used in the algorithm's index, we define the gap $\Delta_i^{\tau,\epsilon}$ between arm $i$ and the threshold $\tau \pm \epsilon$ as

@@

Note that the complexity of the problem increases in the number of arms $K$ and decreases in the gap $\Delta_i^{\tau,\epsilon}$. Consequently, when we require a smaller precision, that is, we increase $\epsilon$, the hardness decreases. Furthermore, we see that each arm contributes its individual complexity, $h_i^{\tau,\epsilon} = \frac{1}{(\Delta_i^{\tau,\epsilon})^2}$ linearly into the problem. Thus, if the complexity $H$ is a good measure of the actual problem complexity, and we knew upfront which arm has which mean, we could allocate the "optimal" share of the budget, $T \cdot h_i /H$ to arm $i$. However, since we do not know neither $\mu_i$ nor $H$, the best we can do is to adapt our allocation of samples more and more towards this optimum as we observe samples.

@@

In this chapter we present the upper bound on the APT algorithm proven in Locatelli et al. (2016). The upper bound is on the probability of error that the algorithm has after $T$ rounds of sampling for a problem with arms that are assumed to be $R$-sub-Gaussian. The bound depends also on the complexity of the problem, and we show in the proof that given the index of the APT algorithm as defined in \autoref{sec:TheAPTAlgorithm}, the complexity $H$ is an appropriate measure.

@@

and similar for $\hat{\Delta}_i$. We can now lower bound the LHS of the indices as

@@

Furthermore, the SLR strategy reduces to the APT strategy in the case where arms are normally distributed with equal variance that is known upfront. In that case we have as index (compare example 3 above): 
$$
B_i^{SLR}(t) = T_i(t)\frac{(\hat{\mu}_i(t) - \tau)^2}{2\sigma^2}.
$$
In order to decide at time $t$ which arm to draw next, we necessarily compare the indices of arm $i$ and arm $j$. Assume w.l.o.g. that $B_i^{SLR}(t) < B_j^{SLR}(t)$. Then:

@@

We compare the empirical performance of the SLR algorithm not only against the APT algorithm, but also against the variance-based algorithms. Thus we also compare the variance-based algorithms for the first time directly against each other. As benchmark, the uniform sampling strategy is included. Additionally, we include the Bayes-UCB algorithm introduced in Kaufmann et al. (2012) for the cumulative regret setting. See \autoref{sec:AppendixBUCB} in the appendix for how their Algorithm 1 is adapted to the thresholding bandit problem.

@@

Every experiment presented is repeated for 5000 iterations. Depending on the experiment, the observations are generated randomly by a given exponential family of distributions. As we are primarily concerned with improving upon APT with respect to non-sub-Gaussian distributions, the first three experiments focus on Bernoulli distributions, especially those with small means. We then go on to show with experiments on data from the Exponential distribution and Poisson distribution that the SLR algorithm is able to improve upon APT given that we know the true kind of distribution.

### Simulation 1: Bernoulli with $\tau = 0.5$

>>>>

The thesis at hand joins a large literature concerned with the problem of sequential sampling. One is familiar with the idea of "batch sampling", where the entire sample is gathered first and analyzed only after the last sample has become available. However, as noted in Chernoff (1958), many situations arise in which one would like to gather feedback from samples as they are drawn, so as to influence the choice of the next sample. The two fundamental characteristics of the latter sampling process are that it is sequential and adaptive. Since observations are drawn in sequence, one has a chance to analyze past observations in order to adapt the decision of what sample one would like to observe next.

This makes sequential sampling especially relevant in situations where there is an inherent cost to obtaining samples. The classical example is that of clinical trials in which a researcher tests two drugs on patients to observe which one performs better. Not only is it difficult to gather enough patients, but one would also like to adapt the sampling such that as soon as possible more patients receive the better drug and are cured. As noted in Cappé et al. (2013), this example dates back to Thompson (1933, 1935). Such problems are known as multi-armed bandit problems in which the sampling is adapted so as to minimize the cumulative regret: In the case of the drug tests, one would like to sample the two drugs in a way to find the best drug as soon as possible and then to test this drug more and more often and more often than the worse drug. This would lead to cumulatively more healthy patients than testing both drugs uniformly; it would minimize the regret from treating patients with an ineffective drug.

Instead of drugs, one can consider two or more unknown distributions (which we call arms in the multi-armed bandit problem in analogy with the one-armed bandit slot machine). These distributions have different means, and one can draw samples sequentially from the distributions in order to estimate their means. To minimize the cumulative regret measure, one would like to play the distribution with the highest mean as much as possible (*Exploitation*). At the same time, however, one has to sample all arms at least a little, so as to initially find the best arm (*Exploration*). Any algorithm thus has to handle a trade-off between exploration and exploitation to minimize the regret. How can we organize the sampling in order to achieve this task?

While Gittins (1979) presents an Bayesian-optimal solution to the cumulative regret problem, the Upper-Confidence Bound (UCB) algorithms pioneered by Agrawal (1995) and Auer, Cesa-Bianchi and Fischer (2002) have lead to a blooming literature on multi-armed bandit problems that has been picked up to solve modern problems in website and advertising optimization, where advertisers try to allocate their advertising optimally so as to reduce regret from showing poorly converting advertisements. Several algorithms like the KL-UCB algorithm based on Kullback-Leibler divergence upper confidence intervals (Cappé et al., 2013) or the Bayes-UCB algorithm (Kaufmann et al., 2012)  have followed and been shown to be optimal as they reach the lower bound introduced in Lai and Robbins (1985) and Burnetas and Katehakis (1996).

These algorithms have in common that they compute after every new sample an index for each of the unknown distributions. This index is based solely on the past (observed) data, and then used to pick the distribution from which the next sample should be drawn. Using these indexes, algorithms can sample sequentially and adapt their strategies to observations as they are collected.

Similar algorithms based on indexes can be used on related problems. Instead of minimizing cumulative regret, it can be of interest to find the best arm, that is, the distribution with the largest mean -- with as few samples as possible and with confidence as large as possible. This is particularly useful when there is not a cost from testing arms with lower mean, but when drawing any sample is costly. Thus one needs to explore the distributions as efficiently as possible. Exploitation is not necessary, and the setting is thus referred to as *Pure Exploration*. The need for strategies other than uniform sampling arises: One does not want to continue to sample distributions that one is confident in. This setting is related to the problem faced in Bayesian Optimization: Training large neural networks is costly in terms of time and computation, and so the parameter space needs to be explored with as few samples as possible. (Zhang et al., 2017; Brochu et al., 2010). The problem of best arm identification is also very related to statistical hypothesis testing as noted by Chernoff (1958) since one has to make a decision about how the means of distributions relate to each other. The main difference being that the bandit setting includes the testing procedure in the sampling process.

Given the focus on exploration, the pure exploration problem is usually described with a loss function different from the one in the cumulative regret settings. Audibert et al. (2010) introduce a soft measure called *simple regret* based on which they describe UCB policies. Other authors investigate binary measures that are equal to 1 when the best arm has not been identified (Russo, 2016; Garivier and Kaufmann, 2016). These binary measures correspond to the probability of the identified arm being the best, and thus inherently related to hypothesis testing. The problem of best arm identification can readily be extended to the so called Top-$m$ problem of identifying the best $m$ of $K$ arms (Kaufmann and Kalyanakrishnan, 2013; Kaufmann et al. 2016).

In this thesis, we consider the thresholding bandit problem. Instead of identifying the best $m$ out of $K$ arms, we aim to find all distributions out of $K$ that have a mean at least as large as a threshold fixed upfront. In the form described in Locatelli et al. (2016), algorithms in the thresholding bandit problem are evaluated against a binary loss function penalizing whenever an arm has not been classified correctly as being above or below the threshold, and thus focus on exploration similar to algorithms employed for best arm identification. Indeed, Locatelli et al. (2016) show how their thresholding algorithm can be used for best arm identification when the mean of the best arm is known upfront.

In the following, we first present the thresholding bandit problem in detail in \autoref{chap:ThresholdingBanditProblem}. To then extend upon the algorithm proposed by Locatelli et al. (2016) and discussed in \autoref{sec:TheAPTAlgorithm}, we first explore in \autoref{chap:KullbackLeiblerPureExploration} how the Kullback-Leibler divergence has been employed in the best arm identification literature by Garivier and Kaufmann (2016) and Kaufmann et al. (2016) as this will lead us to our own extension. We shortly describe in \autoref{sec:VarianceBasedAlgorithms} two variance-based algorithms for the thresholding bandit problem introduced in Mukherjee et al. (2017) and Zhang et al. (2017), before introducing the new Simple Likelihood Ratio algorithm in \autoref{sec:LRforUnivariateExponentialFamily}. We end in \autoref{chap:Experiments} with experiments to examine the performance of different algorithms.

@@

To introduce the thresholding bandit problem, consider a game in which the learner is tasked to find the slot machines within a row of $K$ slot machines in a casino that give a mean payoff larger than a fixed threshold $\tau$, for example $\tau = 0$. The learner cannot pull all slot machines at the same time. Instead, he acquires samples sequentially from one slot machine at a time to learn about their payoff structure. The learner has a budget of $T>K$ slot machine pulls. After $t-1$ samples, the learner has observed $T_i(t-1)$ samples from slot machine $i$. The payoffs $X_{i,1}, ..., X_{i,T_i(t-1)}$ are drawn from the payoff distribution $\nu_i$ underlying the slot machine $i$. At time $t$, the learner considers the payoffs he observed so far in order to choose the next slot machine he would like to pull. Her decision is lead by the goal to classify the slot machines with distribution $\nu_i$ and mean payoff $\mu_i = \mathbb{E}_{X \sim \nu_i}[X]$ after $T$ samples as being above or below the threshold with error probability as small as possible.

The thresholding bandit problem can thus be framed in the literature on sequential sampling strategies. Even though its name is reminiscient of the multi-armed bandit setting, the thresholding bandit problem is more akin to best arm identification and top-$m$ identification problems due to its focus on exploration: While the classical multi-armed bandit problems are concerned with cumulative regret which increases *during* sampling, best arm identification and thresholding bandits collect loss only once *after* the sampling procedure (compare Garivier and Kaufmann, 2016; Kaufmann et al., 2016; Kaufmann and Kalyanakrishnan, 2013). As we will see in \autoref{chap:KullbackLeiblerPureExploration} there arises an implicit threshold between the means of the $m$-th best arm and the $m+1$-th best arm in the top-$m$ identification problem (Kaufmann et al., 2016; Kaufmann and Kalyanakrishnan, 2013). Because of this, the two settings are very related.

@@

Consider a setting with a set of $\mathbb{A}=\{1, ..., K\}$ unknown probability distributions $\nu_i, i \in \mathbb{A}$. To stick with the idea of multi-armed bandits, we call the distributions arms, and observations from the arms rewards (Bubeck and Cesa-Bianchi, 2012). The learner is interested in learning about the structure of the arms; in particular, she would like to estimate the means $\mu_1, ..., \mu_K$ by sequentially drawing observations $X_{i,1}, X_{i,2}, ...$ from the different arms. However, at time $t$ the learner cannot observe a reward from every distribution but must select a single arm $I_t$ from which she would like to observe the reward $X_{I_t,t}$. Often, adaptive sampling strategies makes the decision of which arm to pull by computing an index $B_i(t)$ for each arm. If $T_i(t)$ is the number of samples observed until time $t$ for arm $i$, then is $B_i(t)$ a function of the past rewards $X_{i,1}, ..., X_{i,T_i(t)}$ observed for arm $i$, and w.l.o.g. $I_t = \arg \min_{i = 1,...K} B_i(t)$. The set of indexes can be called strategy (Kaufmann et al., 2012) which makes it clear that the strategy is adapted as more observations are collected. Strategies of this kind are also called index policies.

Which index proves to be useful is determined by the loss function with which the decisions of the algorithm are evaluated. Many problems are concerned with finding the arm with the largest mean $\mu_i^*$ while collecting loss either cumulatively during (Auer and Cesa-Bianchi, 2012) or once at the end of the sampling (Kaufmann et al., 2016). The thresholding bandit problem considers a slightly different loss function. Consider a fixed threshold value $\tau \in \mathbb{R}$ and an epsilon interval $\tau \pm \epsilon$ around the threshold for $\epsilon \geq 0$. As mentioned in the introduction, the learner is tasked with classifying arms as having a mean $\mu_i$ larger or smaller than $\tau$. As in Locatelli et al. (2016), whenever we say that "an arm is over the threshold", we say that the mean of the arm is over the threshold. To describe which arms are over the threshold, we also introduce the following set notation: Let $\mathcal{S}_{u}$ be the set of distributions in $\mathbb{A}$ with mean larger than some $u \in \mathbb{R}$: $\mathcal{S}_{u} := \{i \in \mathbb{A}, \mu_i \geq u\}$. Consequently, the complement is given by the arms with mean below $u$, $\mathcal{S}^C_{u} := \{i \in \mathbb{A}, \mu_i < u\}$.

The thresholding problem for the learner is now to use a budget of overall $T$ samples (or rounds) to explore the different arms. After this exploration phase, the learner returns a set of arms that she believes to be over the threshold. When evaluating the classification of the learner, arms with mean inside of the epsilon interval $\tau \pm \epsilon$ will not be considered. The learner only has to correctly distinguish the means from the threshold up to the precision given by $\epsilon$. Thus, the algorithm is not evaluated against $\mathcal{S}_{\tau}$ and $\mathcal{S}^C_{\tau}$, but needs to distinguish $\mathcal{S}_{\tau+\epsilon}$ from $\mathcal{S}^C_{\tau-\epsilon}$.

@@

$$
\mathcal{L}(T) = \mathbbm{1}{\big((\hat{\mathcal{S}}_{\tau} \cap \mathcal{S}_{\tau-\epsilon}^C \neq \emptyset) \lor (\hat{\mathcal{S}}_{\tau}^C \cap \mathcal{S}_{\tau+\epsilon} \neq \emptyset)\big)}
$$

In order to perform well, a learner has to identify all arms that are over the threshold as such, and to not claim arms as being over the threshold if they actually are below the threshold. Due to the precision $\epsilon$, the loss is not influenced by arms which are very close to the threshold $\tau$ as measured by $\epsilon$, that is, arms inside of the $\tau \pm \epsilon$ interval. If either $\hat{\mathcal{S}}_{\tau}$ contains an arm below $(\tau - \epsilon)$, or $\hat{\mathcal{S}}^C_{\tau}$ contains an arm above $(\tau + \epsilon)$, a loss of $1$ is accrued. Else, all arms outside the epsilon interval are classified correctly and the loss is equal to $0$. In the case of $\epsilon = 0$, all arms need to be classified correctly to not collect any loss.

@@

which corresponds to the probability of making at least one wrong classification. Both the probability of not identifying an arm that is actually over the threshold, as well as the probability of accepting an arm as over the threshold when it is not, are evaluated. The oracle strategy would evaluate all arms correctly with an error probability of $0$ (Locatelli et al., 2016).

At this point, we have introduced the general thresholding bandit problem setting which we will use throughout the thesis. The setting can be extended by different assumptions to make certain algorithms or bounds feasible. For that reason, Locatelli et al. (2016) assume that the distributions $\nu_i, i \in \mathbb{A}$ are $R$-sub-Gaussian. Under this assumption, the authors are able to prove a lower bound on the largest probability of error of the best algorithm in the setting we have just introduced, as well as an upper bound on the performance of their APT algorithm which we will introduce in the following chapter. As in their paper, we define $R$-sub-Gaussian and assume it in the description of the their algorithm and bounds:

@@

As noted by Locatelli et al. (2016), this ensures that the mean $\mu_i$ of the distributions $\nu_i$ considered in our problem are finite. $R$-sub-Gaussian distributions offer concentration inequalities for the concentration of the empirical mean around the true mean, which makes it possible to deduce the bounds presented by the authors and repeated below. The family of $R$-sub-Gaussian distributions encompasses Gaussian distributions of variance $R^2$ for $R \in \mathbb{R}$, as well as all bounded distributions, with the important case of Bernoulli distributions.

@@

Just like the general index policies mentioned in \autoref{sec:Setup}, we see that the APT algorithm adapts its strategy using an index $B_i(t+1)$ for each arm that is based entirely on the observations observed until time $t$. Given the arm with the smallest index is pulled at round $t+1$, we observe that arms that are estimated to be further away from the threshold $\tau$ as measured by the absolute distance to their empirical mean $\hat{\mu}_i$, $\hat{\Delta}_i(t)$, are pulled less often. Intuitively, it is easier to classify arms as below or above $\tau$ when their mean is far away from the threshold. In contrast to the uniform sampling strategy, APT will gather more evidence for arms which appear to be inherently more difficult to classify. Likewise, APT tends to less often pull arms which it already pulled considerably often as measured by $T_i(t)$. If arm $i$ has been pulled more often than arm $j$, then we should in general be more confident in its estimated mean and the corresponding distance to the threshold. If we are confident in $i$, then it is better to explore other arms by drawing an observation from arm $j$. By adapting its sampling strategy in this way during the sequential sampling process, the APT algorithm minimizes the probability of error $\mathbb{E}[\mathcal{L}(T)]$ after $T$ samples have been observed.

The previous paragraph describes how arms close to the threshold are intuitively more difficult to classify. In the index employed by the APT algorithm, the distance between an arm and the threshold is measured by the absolute distance $\hat{\Delta}_i(t)$ between the (estimated) mean of the arm and the threshold $\tau$ (and the epsilon interval). When the APT algorithm realizes that an arm has a mean close to the threshold, then this arm requires more attention than an arm far away from the threshold; the confidence in the empirical mean has to be larger. Or the empirical mean has to be better concentrated around the true mean. Otherwise it is easier to make a mistake on this arm. To gain this confidence, the arm requires more samples, and thus a larger share of the overall $T$ samples available. With this intuition, it makes sense to measure the complexity of a thresholding problem: If we knew upfront how complex a given set of arms is, we would know whether to expect a small or large probability of error given our budget $T$. Also, for example in the best arm identification problem, there exist algorithms which take the complexity of a problem as input. Based on the complexity they then adapt their sampling strategy. See for example the UCB-E algorithm in Audibert et al. (2010) for the best arm identification problem, or the KL-LUCB-E algorithm in Kaufmann and Kalyanakrishnan (2013) as extension to the top-$m$ problem.

While the APT algorithm does not require the complexity to be known upfront (which would be problematic as it is not known upfront), it is necessary to define the complexity to deduce the upper bound on its probability of error. Similar to the distance measure used in the algorithm's index, we define the gap $\Delta_i^{\tau,\epsilon}$ between arm $i$ and the threshold $\tau \pm \epsilon$ as

@@

Note that the complexity of the problem increases in the number of arms $K$ and decreases in the gap $\Delta_i^{\tau,\epsilon}$. Consequently, when we require less precision, that is, we increase $\epsilon$, the hardness decreases. Furthermore, we see that each arm contributes its individual complexity, $h_i^{\tau,\epsilon} = \frac{1}{(\Delta_i^{\tau,\epsilon})^2}$ linearly into the problem. Thus, if the complexity $H$ is a good measure of the actual problem complexity, and we knew upfront which arm has which mean, we could allocate the "optimal" share of the budget, $T \cdot h_i /H$ to arm $i$. However, since we do not know neither $\mu_i$ nor $H$ upfront, the best we can do is to adapt our allocation of samples more and more towards this optimum as we observe samples.

@@

In this chapter we present the upper bound on the probability of error that the APT algorithm incurs, as proven in Locatelli et al. (2016). The upper bound is on the probability of error that the algorithm has after $T$ rounds of sampling for a problem with arms that are assumed to be $R$-sub-Gaussian. The bound depends also on the complexity of the problem, and we show in the proof that given the index of the APT algorithm as defined in \autoref{sec:TheAPTAlgorithm}, the complexity $H$ is an appropriate measure.

@@

and similar for $\hat{\Delta}_i$. We can now lower bound the LHS of the indexes as

@@

Furthermore, the SLR strategy reduces to the APT strategy in the case where arms are normally distributed with equal variance that is known upfront. In that case we have as index (compare example 3 above): 
$$
B_i^{SLR}(t) = T_i(t)\frac{(\hat{\mu}_i(t) - \tau)^2}{2\sigma^2}.
$$
In order to decide at time $t$ which arm to draw next, we necessarily compare the indexes of arm $i$ and arm $j$. Assume w.l.o.g. that $B_i^{SLR}(t) < B_j^{SLR}(t)$. Then:

@@

We compare the empirical performance of the SLR algorithm not only against the APT algorithm, but also against the variance-based algorithms. Thus we also compare the variance-based algorithms for the first time directly against each other. As benchmark, the uniform sampling strategy is included. Additionally, we include the Bayes-UCB algorithm introduced in Kaufmann et al. (2012) for the cumulative regret setting. See \autoref{sec:AppendixBUCB} in the appendix for how we adapt the Bayes-UCB algorithm to the thresholding bandit problem.

@@

Each of the following five simulations is repeated for 5000 iterations. Depending on the simulation, the observations are generated randomly by a given exponential family of distributions. As we are primarily concerned with improving upon APT with respect to non-sub-Gaussian distributions, the first three simulations focus on Bernoulli distributions, especially those with small means. We then go on to show with experiments on data from the Exponential distribution and Poisson distribution that the SLR strategy can be adapted to other distributions as well.

### Simulation 1: Bernoulli Arms and $\tau = 0.5$
